---
title: "Lesson 4 HW 2 EC"
author: "Ellen Chancey"
date: "September 16, 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fmsb)
library(PropCIs)
```

###Question 3.3 a and c  

####Part A  
```{r}
or<- oddsratio(60,2,44,61,conf.level = 0.95)
#exposed is political positon and diseased is agree
logOR <- log(or$estimate)
logOR
```

####Part C
```{r}
# library(fmsb)
# Estimate, confidence interval, and p value for odds ratio
oddsratio(60,2,44,61,conf.level = 0.95)
```


```{r}
# reccomended package doesn't come up with what the book says
#library(PropCIs)
orscoreci(60,(60+44),2,(2+61),conf.level=0.95)

```

The interval is imprecise because there are very few numbers of republicans who agreed.  



###Question 3.4

difference of proportions confidence interval
```{r}
wald2ci(1085,(1085+55623),703,(703+441239), conf.level=0.95, adjust = "Wald")
```
The difference of proportions confidence interval indicates that there is a difference in the proprotion of fatalities between those who do and do not wear a seatbelt.  
The confidence level indicates that 95% of the time this method will produe a range that reflects the true parameter, in this case difference of proportion. In 5% of cases this methodology will produce an interval that does not reflect the true parameter.  


relative risk confidence interval
```{r}
riskscoreci(1085,(1085+55623),703,(703+441239), conf.level=0.95)
```
The confidence interval of the relative risk indicates that the risk of fatality when not wearing a seatbelt is relatively higher than when wearing a seatbelt.
The confidence level indicates that 95% of the time this method will produe a range that reflects the true parameter, in this case relative risk. In 5% of cases this methodology will produce an interval that does not reflect the true parameter.  


oddsratio confidence interval
```{r}
#library(fmsb)
oddsratio(1085,703,55623,441239,conf.level=0.95)
```
The odds ratio confidence interval indicates that those not wearing a seatbelt are 11.13 to 13.47 times more likely to experience a fatality.
The confidence level indicates that 95% of the time this method will produe a range that reflects the true parameter, in this case the odds ratio In 5% of cases this methodology will produce an interval that does not reflect the true parameter.




```{r}
sessionInfo()
```

