---
title: "Week 1 HW Ellen Chancey"
author: "Ellen Chancey"
date: "August 25, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1.2 
100 questions on a multiple choice test. There are 4 options for each question, one is correct. A student randomly guesses on each question.  
There are 100 trials, each with 2 possible outcomes - right or wrong. Each trial has a 0.25 probablity that the outcome is "correct".

A. Specify the distribution of the number of correct answers  
The distribution of outcomes is binomial because the answer is either wrong or right for each trial. Across 100 trials, there can be between 0 and 100 correct answers. 

B. Find the mean and standard deviation of that distribtuion. Would it be surprising if the student made at least 50 correct responses? Why?  
The mean of the binomial distribution is the \(E(y) = \pi n\).
```{r}
n <- 100 # trials
p <- 0.25 # probablity that correct outcome is selected for one trial
c <- 4 # number of outcomes
mu <- p*n
mu
```

The variance of the binomial distribution is \(\pi(1-\pi)\)
```{r}
var <- n*p*(1-p)
var
```

It would be surprising for a student to obtain 50 correct response becase the mean + the standard deviation is 43.75. It is unlikely a student would obtain 50 correct answers given these statistics.
```{r}
low <- mu - var
low
high <- mu + var
high
```


C. Specify the distribution of \((n_1,n_2,n_3,n_4)\) where \(n_j\) is the number of times the student picked choice \(j\).  

```{r}

```

D. Find \(E(n_j)\) and \(var(n_j\). Show that \(cov(n_j,n_k) = -6.25\) and \(corr(n_j,n_k) = -0.333\)  
```{r}

```



### Question 1.4  
Russian roulette, where a bullet is placed in one of six chambers of a pistol, the chamber is spun and a random chamber is selected at random.  

A. What is the probability of doing this six times where none of the trails results in a bullet firing?  
The probability that in six trials, it never fires is 0.3348
```{r}
n <- 6 # number of trials
y <- 0 # number of fires
p <- 1/6 # probability that is fires in one trial
prob <- (choose(n,y))*(p^y)*((1-p)^n-y)
prob
```


### Supplemental Question  
Find \(E(Y)\) and \(Var(Y)\) when \(Y=Poi(\mu)\)






Session Info
```{r}
sessionInfo()
```