---
title: "Lecture 11 HW"
author: "Ellen Chancey"
date: "September 29, 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Question 3.19  
A study in the department of wildlife ecology sampled wild common carp fish from a wetland in central Chile. One analysis investigated whether the fish muscle had lead pollutant and whether there was evidenct malformation in the fish. Of 25 fish without lead, 7 had malformation. Of 14 with lead, 7 had malformation. Report and interpret the p value for Fisher's exact test for a one sided alternative of a greater chance of malformation when there is lead pollution.
```{r}
fish <- matrix(c(7,7,7,18), byrow = TRUE, nrow = 2)
dimnames(fish) <- list(lead=c("yes", "no"), 
                       malform=c("yes", "no"))
fish
```


```{r}
fisher.test(fish, alternative = "greater", conf.level = 0.95)
```
Indicates that the probability of a table with a larger number of malformation in fish with lead is 0.1526.  



### Additional Problem  
A contingency table for two independent binomial variables where Y=1 denotes success and Y=2 failure has three of three successes in row one and zero of three successes in row two.  For H0: Pi_1<=Pi_2  versus Pi_1>Pi_2 , show that the p-value for the unconditional exact test equals 1/64 and for Fisher’s exact test equal 1/20.  (Hint: see p 94, which has H0: Pi_1=Pi_2.)
```{r}
ex <- matrix(c(3,0,0,3), byrow = TRUE, nrow = 2)
dimnames(ex) <- list(x=c("a","b"),
                     y=c("success","fail"))
ex
```

Unconditional Exact Test
```{r}
library("Exact")
exact.test(ex, alternative="greater")
```


Fisher's Test
```{r}
fisher.test(ex,alternative = "greater", conf.level = 0.95)
```












### Session Info
```{r}
sessionInfo()
```

