---
title: "Lesson 5 HW 1 Chancey"
author: "Ellen Chancey"
date: "September 22, 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
 # My "procfreq" function: 
"procfreq" <- function(x, digits=4) 
        {
          total <- sum(x)
          rowsum <- apply(x,1,sum)
          colsum <- apply(x,2,sum)
          prop <- x/total
          rowprop <- sweep(x,1,rowsum,"/")
          colprop <- sweep(x,2,colsum,"/")
          expected <- (matrix(rowsum) %*% t(matrix(colsum))) / total
          dimnames(expected) <- dimnames(x)
          resid <- (x-expected)/sqrt(expected)
          adj.resid <- resid /
          sqrt((1-matrix(rowsum)/total) %*% t(1-matrix(colsum)/total))
          df <- prod(dim(x)-1)
          X2 <- sum(resid^2)
          attr(X2,"P-value") <- 1-pchisq(X2,df)
          tmp <- x*log(x/expected)
          tmp[x==0] <- 0
          G2 <- 2 * sum(tmp)
          attr(G2,"P-value") <- 1-pchisq(G2,df)
          list(sample.size=total,
          row.totals=rowsum,
          col.totals=colsum,
          overall.proportions=prop,
          row.proportions=rowprop,
          col.proportions=colprop,
          expected.freqs=expected,
          residuals=resid,
          adjusted.residuals=adj.resid,
          chi.square=X2,
          likelihood.ratio.stat=G2,
          df=df)
        }

```


#### 3.11 parts a and b  

Set Up
```{r}
raceparty <- matrix(c(192,75,8,459,586,471), byrow = TRUE, nrow = 2)
dimnames(raceparty) <- list(race=c("Black", "White"), 
                            party=c("Democrat", "Independent", "Republican"))
raceparty
```

G Test
```{r}
library("DescTools")
rpg <- GTest(raceparty)
rpg
```
The \(G^2\) value of this contingency table is very high. This provides evidence to reject the null hypothesis, \(H_0: \pi_{ij} = \pi_{i+} \pi_{+j} \). The p-value indicates that this method will rarely reject the \(H_0\) when its true.  


Standardized Residuals
```{r}
rppf<- procfreq(raceparty)
rppf$adjusted.residuals
```


### Additional Problem 1  
Investigators studied the relationship between stage of breast cancer at diagnosis (local or advance) and a woman’s living arrangement.  Of 144 women living alone, 41.0% had an advanced case; of 209 living with a spouse, 52.2% were advanced; and of 89 living with others, 59.6% were advanced.  The investigators reported their P-value for the relationship as 0.02 (Moritz and Satariano, J. Clin. Epidemiol. 45:443-454, 1993).  Reconstruct the analysis performed to obtain this P-value.

Set Up
```{r}
cancer <- matrix(c(
            round(144*.41),
            round(209*.522),
            round(89*.596),
            round(144*.59),
            round(209*.478),
            round(89*.404)),
            byrow = TRUE, 
            nrow = 2)
dimnames(cancer) <- list(stage=c("Advanced", "Local"), 
                        living=c("Alone", "Spouse", "Others"))
cancer <- t(cancer)
cancer
```

Method One
```{r}
library(MASS)
chisq.test(cancer)
```

Method Two
```{r}
cancerpf <- procfreq(cancer)
cancerpf$chi.square
```

I assume that researchers round up, given that I obtain a p value of 0.015.  

### Additional Problem 2  
Table 3.13 (p 106) presents a contingency table for a sample of patients by their diagnosis and whether they were prescribed medication.  Obtain standardized Pearson residuals for the independence model.  Interpret these results.

Set Up
```{r}
drug <- matrix(c(105,8,12,2,18,19,47,52,0,13), byrow = TRUE, nrow = 5)
dimnames(drug) <- list(diagnosis=c("Schizophrenia", "Affective", "Neurosis", "Personality", "Special"), 
                            drug=c("Yes", "No"))
drug
```


```{r}
chisq.test(drug, correct = TRUE)
```
The \(X^2\) value is 84.188 with a very small p value, well below 0.05. This result provides evidence to reject the null hypothesis - \(H_0: \pi_{ij} = \pi_{i+} \pi_{+j} \). The p-value indicates that this method will rarely reject the \(H_0\) when its true.  



#### Session Info
```{r}
sessionInfo()
```

