---
title: "Lesson 2 Homework Ellen Chancey"
author: "Ellen Chancey"
date: "August 31, 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Question 1.5  
People are surveyed with whether they think it should be possible for a pregnant woman to obtain a legal abortion if she is married and does not want any more children.
```{r}
yes <- 587
no <- 636
n <- yes + no
pi <- yes/n
```



####Find the p-value for \(H_0: \pi = 0.50\) and construct 95% confidence interval

```{r}
binom.test(yes,n, p=0.5, conf.level = .95, alternative = "two.sided")
```


####Interpret the results  
The confidence interval includes 0.50 and the p value is large, therefore the null hypothesis that \(\pi = 0.50\) cannot be rejected.




## Additional Question
A sample of 100 women who suffer from dysmenorrhea participated in a study.  A new analgesic is claimed to provide greater relief than a standard one.  After using each analgesic in a crossover experiment, 40 of the participants reported greater relief with the standard analgesic and 60 reported greater relief with the new analgesic.  Analyze these data.
```{r}
total <- 100
standard <- 40
new <- 60
pi2 <- new/total
```

Null hypothesis test and confidence interval  
\(H_0 : \pi = 0.50\)  
\(H_1 : \pi <> 0.50\)
```{r}
binom.test(new,total, p=0.5, conf.level = .95, alternative = "two.sided")
```

####Interpret the results 
This is a very border line result. A conservative approach would indicate that the null hypothesis cannot be rejected because the p value is more than 0.05 and because the confidence interval contains the value indicated in the null hyothesis.  


####Session Info
```{r}
sessionInfo()
```

